---
title: "Quarto Set Up, Descriptive Statistics and Hypotheses Testing"
shorttitle: "Quarto Descriptive Hypotheses"
author:
  - name: Yimeng (Amy) CHENG
    corresponding: true
    orcid: 0009-0007-1238-3749
    email: yimengcheng@uchicago.edu
    affiliations:
      - name: University of Chicago
        department: Department of Psychology
        address: 5801 S Ellis Ave
        city: Chicago
        region: IL
        country: USA
        postal-code: 60615
author-note:
  status-changes: 
    affiliation-change: null
    deceased: null
abstract: "This assessment uses the math anxiety dataset to perform various statistical analyses and hypotheses testing."
keywords: ["Math Anxiety", "Cognitive Concerns", "Perfectionism", "Statistical Anxiety"]
bibliography: "bibliography.bib"


format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{listings}
        \lstset{
          breaklines=true
        }
---

```{r}
#| label: set up
#| include: FALSE
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
```

# Objective-modified

This assessment is based on '04_data-analysis/01_descriptives', but it has been greatly modified to satisfy assessed and unassessed goals in the following areas:

1.  Basic R set up.
2.  Quarto set up.
3.  APA formatting and reference.
4.  Creating figures using ggplot2.
5.  Perform statistical analyses, including descriptive and hypotheses testing, and present the results in a proper way.
6.  Render to PDF.

The dataset being used in the assessment is 'math-anxiety.csv' from Natalie's '04_data-analysis' folder. You may also notice that a lot of instructions are also from the original '04_data-analysis/01_descriptive/descriptive-stats.qmd' file. That is because I was originally doing this task but then thought I could play around with this task and meet extra unique goals. However, the end product did turnout to be quite messy and I apologize for that :( I promise that I will do better in my final research project.

*Note:* To meet APA formatting requirements, in the **PDF** rendered, **only the proper APA tables and figures will be presented**, and they are at the very end of the document. However, the codes for the walkthrough process and intermediate is figures still presented.

# Objective-original

The objective of this assignment is to practice using R to calculate and interpret descriptive statistics. Unlike in previous assignments, you will not use a built-in dataset. Instead, you will use data from a published, open-access dataset capturing relationships between math anxiety and self-perception:

*Mackinnon, S. P., McCaughey, N., & Hill, T. G. (2024, July 25). The Association of Self-Efficacy, Anxiety Sensitivity, and Self-Critical Perfectionism with Statistics and Math Anxiety. https://doi.org/10.17605/OSF.IO/NZHQ6*

You can access the full dataset and codebook [here](https://osf.io/nzhq6/). The data you will need is included in the .csv file within this project's directory.

Before you begin diving into the data summary, you should take some time to familiarize yourself with the dataset, including the different categories of variables and how they can be interpreted.

# Background information

It is widely recognized that anxiety is a common problem among university students[@leviness2017]. However, fewer people know that besides general anxiety, domain-specific anxieties are also prevalent. These includes math anxiety , statistical anxiety and so forth [@onwuegbuzie2003].

The current project investigates the relationships among domain-specific anxieties and its related constructs, as well as some demographic variables among university students. The dataset is from @mccaughey2022.

# Set Up

```{r}
#| label: load library

library(tidyverse)
library(tidyr)
library(purrr)
library(tibble)
library(tidyr)
library(knitr)
library(apaTables)
library(ggplot2)
library(kableExtra)
library(forcats)

set.seed(1234)

## I also used quarto::quarto_use_template("wjschne/apaquarto", no_prompt = TRUE) to install quartoand its apa extension for RStudio
```

## Read in the data

Run this chunk to read in the dataset and view the first few rows:

```{r}
#| label: load-goal-tibble
#| echo: TRUE
#| results: hide

math.anxiety <- read_csv("math-anxiety.csv") 
```

```{r}
#| label: goal-tibble
#| echo: TRUE
#| results: hide

# View first 6 rows
head(math.anxiety)

```

Examine the structure of the df. Take note of data types and review factor levels.

```{r}
#| label: examine-structure
#| echo: TRUE
#| results: hide

# Examine the structure of the df
str(math.anxiety)

# Seems like there are only two types of data: numbers and characters.

#change character data to factor so that I can see different levels
math.anxiety <- math.anxiety %>%
  mutate(across(where(is.character), as.factor))
  
math.anxiety %>%
  select(where(is.factor)) %>%
  map(levels) 

```

## Data Preparation

Not all variables are currently the data type we'd like them to be. Based on what you see when you look at the structure of the df and using the data's [codebook](https://osf.io/zh5ab), which columns are not currently in the correct data type? What changes do you need to make?

<!-- answer below -->

<!-- Hint: There are a lot of columns, and the vast majority are numeric. start by looking through the codebook to figure out what *shouldn't* be numeric and go from there. -->

**Answer:** Program type should not be numeric but should be factor. Id should also not be numeric but should be character.

In the chunk below, perform simple data cleaning to retype the columns as needed.

```{r}
#| label: data-type-cleaning
#| echo: TRUE
#| results: hide

# Change the data type of columns as needed
math.anxiety <- math.anxiety %>%
   mutate(program.type = factor(program.type, levels = c(0, 1), labels = c("Undergraduate", "Graduate")), id = as.character(id))

# Change program.type to factor, with two levels: 1 prepersents yes and 0 represents no.
#Change id to character so that it would not be calculated in summary.


```

In an effort to simplify things at least a little, let's drop the columns that are item responses, keeping participant info and composite survey scores. Use the codebook to identify which columns are item responses, then modify the dataframe to drop those columns. We're assigning this to a new object, `math.anx`, to keep the original data intact.

```{r}
#| label: drop-item-responses
#| echo: TRUE
#| results: hide

# There are (at least) two ways to approach this, either:
# 1. Notice that all the item responses follow a naming pattern, then use select helpers to drop them
# 2. Use the codebook to identify the columns that are *not* item responses and keep them

# Assign your new long df to math.anx
math.anx <- math.anxiety %>%
  select(id, age, gender.category, ethnicity.r, program.type, major, dal.course.code, faculty, stats, stats.history, self.efficacy, asi.phys, asi.cog, asi.social, asi, perf.cog, perf.sp, frost.ps, frost.com, frost.da, stat.anx.tc, stat.anx.i, stat.anx.ah, stat.anx.ws, stat.anx.fst, stat.anx.sc, math.anx)


```

# Descriptive Statistics

## Center and Spread

We can see basic summary statistics from the wide data (`math.anx`):

```{r}
#| label: summary-stats
#| echo: TRUE
#| results: hide

# Use the appropriate function to see basic summary statistics for the wide data

summary(math.anx)


```

But that's kind of a mess, and it doesn't let us use the summary statistics for any purpose beyond just looking at them. We can use the long version to calculate summary statistics for each survey score.

The dataset is not currently in a tidy format, at least not for our purposes. The survey scores are spread across multiple columns. If we want to use a survey score as a grouping variable or calculate summary statistics, we need to reshape the data so that each survey score is in a single column.

Now pivot the data so that the survey scores are in a single column. You'll want to retain the participant information in the long format as well.

```{r}
#| label: pivot-longer
#| echo: TRUE
#| results: hide

# Create a list of the survey score variables


# Use pivot_longer to reshape the data so that the survey scores are in a single column


math.anx <- math.anx %>%
  pivot_longer(
    cols = c(asi.phys, asi.cog, asi.social, asi, perf.cog, perf.sp, 
             frost.ps, frost.com, frost.da, stat.anx.tc, stat.anx.i, 
             stat.anx.ah, stat.anx.ws, stat.anx.fst, stat.anx.sc, math.anx), 
    names_to = "Survey_Type",   # Create a new column named “survey_type"
    values_to = "Score"         # Put survey values to "score"
  )

head(math.anx)

```

With the data in a tidy format, we can both calculate summary statistics and visualize the data more easily. Let's start with summarizing mean, median, sd, min/max, and range:

```{r}
#| label: summarize-survey-scores
#| echo: TRUE
#| results: hide

# Use the long data to calculate summary statistics for each survey score

summary(math.anx) #this is the most dirct way

#I can also try a more sophisticated way:

# Calculate summary statistics for each survey score
survey_summary <- math.anx %>%
  group_by(Survey_Type) %>%  # Group by survey type
  summarise(
    Mean = mean(Score, na.rm = TRUE),  # Mean value
    Median = median(Score, na.rm = TRUE),  # Median value
    SD = sd(Score, na.rm = TRUE),  # Standard deviation
    Min = min(Score, na.rm = TRUE),  # Minimum value
    Max = max(Score, na.rm = TRUE),  # Maximum value
    Range = Max - Min  # Calculate range
  )


```

There's a problem with the summary above. Missing values for some scores for some participants mean that the summary calculations are also missing. In some cases that may be what we want, but in this case we have enough data that we can just ignore missing values.

**For some reason I just could not load flextable. I tried search on Google and ask AI as well as friends but nothing worked so far. For now, I will use kableExtra instead to create APA formatted table.**

```{r}
#| label: create-apa-summary
#| echo: TRUE
#| results: hide

# Rename survey types for plain English labels
math.anx <- math.anx %>%
  mutate(Survey_Type = recode(Survey_Type,
    "math.anx" = "Math Anxiety",
    "asi" = "Global ASI",
    "asi.phys" = "Physical concerns",
    "asi.cog" = "Cognitive concerns",
    "asi.social" = "Social concerns",
    "perf.cog" = "Perfectionism Cognitions",
    "perf.sp" = "Perfectionism Self-Presentation",
    "frost.ps" = "Perfectionism Personal Standard",
    "frost.com" = "Concern Over Mistakes",
    "frost.da" = "Doubt about Actions",
    "stat.anx.tc" = "Test and Class Anxiety",
    "stat.anx.i" = "Interpretation",
    "stat.anx.ah" = "Asking for help",
    "stat.anx.ws" = "Worth of Statistics",
    "stat.anx.fst" = "Fear of Statistics",
    "stat.anx.sc" = "Computational Self-concept"
  ))

# Summarize survey scores

survey_summary <- math.anx %>%
  group_by(Survey_Type) %>%
  summarise(
    Mean = round(mean(Score, na.rm = TRUE), 2),
    Median = round(median(Score, na.rm = TRUE), 2),
    SD = round(sd(Score, na.rm = TRUE), 2),
    Min = round(min(Score, na.rm = TRUE),0),
    Max = round(max(Score, na.rm = TRUE),0),
    Range = Max - Min
  ) %>%
  rename("Survey Type" = Survey_Type
  )
```

Because I will need to create APA table multiple times, I will just define a function to simplify the process. The function includes a few essential lines to produce an APA table, but additional lines are needed to talior the table to specific needs:

```{r}
#| label: define kable function for apa formatting
#| echo: TRUE
#| results: hide


kable_apa <- function(summary_data) {
  summary_data %>%
    kable("latex", booktabs = TRUE) %>%
    kable_styling(latex_options = c("striped", "scale_down"), font_size = 10, , position = "float_left")
}

```

```{r}
#| label: tbl-apa-summary-table
#| tbl-cap: "Summary Statistics for Survey Scores"
#| echo: TRUE


# Convert to APA formatted table

kable_apa(survey_summary) %>%
  add_header_above(c(" " = 1, "Descriptive Statistics" = 6)) %>%
  column_spec(1, bold = TRUE)

```

As can be observed from @tbl-apa-summary-table, for most variables, the median and mean are close to each other, indicating a relatively symmetric distribution.

We can also visualize the data to get a better sense of the distribution of scores. Let's start with a histogram of the math anxiety scores.

```{r}
#| label: fig-histogram-math-anxiety
#| fig-cap: "Distribution of Math Anxiety Scores"
#| echo: TRUE


math_anxiety_data <- math.anx %>%
  filter(Survey_Type == "Math Anxiety")

ggplot(math_anxiety_data, aes(x = Score)) +
  geom_histogram(binwidth = 0.5, fill = "steelblue", color = "black", size = 1.2, alpha = 0.7) +
  labs(
    title = "Distribution of Math Anxiety Scores",
    x = "Math Anxiety Score",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
  )

```

### Extra step: Math Anxiety vs. Statistics Anxiety Scores

```{r}
#| label: fig-math-stats-anxiety
#| echo: TRUE

ggplot(math.anx %>% filter(Survey_Type %in% c("Math Anxiety", "Test and Class Anxiety")), aes(x = Score, fill = Survey_Type)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.5, color = "black") +
  labs(
    title = "Comparison of Math and Statistics Anxiety Scores",
    x = "Score",
    y = "Frequency",
    fill = "Survey Type"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
```

We can observe from @fig-histogram-math-anxiety and @fig-math-stats-anxiety that math anxiety scores are relatively normally distributed with a slight positive skew, and test and class anxiety scores are negatively skewed.

I can also apply facet_wrap() and facet_grid() functions to compare the math anxiety and Test and Class Anxiety scores by program type:

```{r}
#| label: fig-math-stat-anx-by-program

ggplot(math.anx %>% filter(Survey_Type %in% c("Math Anxiety", "Test and Class Anxiety")), 
       aes(x = Score, fill = Survey_Type)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.5, color = "black") +
  facet_grid(program.type ~ Survey_Type) +
  labs(
    title = "Comparison of Math and Statistics Anxiety Scores by Program Type",
    x = "Score",
    y = "Frequency",
    fill = "Survey Type"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

```

@fig-math-stat-anx-by-program shows that for both graduates and undergraduates, the math anxiety score is slightly positively skewed, where as text and class anxiety is negatively skewed.

Now generate a violin plot of all the survey scores to see differences in distributions across surveys.

**Creating publication quality plots using theme and labs layers**

```{r}
#| label: fig-violin-survey-scores-publication-quality
#| fig-cap: "Distributions of Survey Scores (Publication Quality)"
#| echo: TRUE

# Create violin plot
ggplot(math.anx, aes(x = Survey_Type, y = Score, fill = Survey_Type)) +
  geom_violin(trim = FALSE, alpha = 0.6, color = "black", size = 1.2) +  # Static aesthetic: black outline, thick borders
  geom_boxplot(width = 0.1, outlier.shape = NA, color = "black", alpha = 0.7) +  # Add boxplot for summary stats
  labs(
    title = "Distribution of Anxiety Scores Across Surveys",
    subtitle = "Comparing different types of anxiety and performance-related scores",
    x = "Survey Type",
    y = "Score",
    fill = "Survey Type"
  ) +
  theme_classic() +  # Professional minimalist theme
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Centered bold title
    plot.subtitle = element_text(size = 12, hjust = 0.2, color = "gray40"),  # Subtle subtitle
    axis.title.x = element_text(size = 14, face = "bold"),  # X-axis label bold
    axis.title.y = element_text(size = 14, face = "bold"),  # Y-axis label bold
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    axis.text.y = element_text(size = 12),  # Readable y-axis labels
    legend.position = "none",  # Remove legend since labels are already clear
     plot.margin = margin(10, 10, 10, 50) #Increase margin so that all labels can be seen 
  )

```

@fig-violin-survey-scores-publication-quality tells us that Perfectionism Self−Presentation is quite negatively skewed, potentially suggesting the presence of outliers or extreme values.

Next, I will analyze all survey scores by program type.

```{r}
#| label: Compute survey scores by program type
#| echo: TRUE
#| results: hide

summary_stats <- math.anx %>%
  group_by(program.type, Survey_Type) %>%
  summarise(
    Mean = round(mean(Score, na.rm = TRUE), 2),
    Median = round(median(Score, na.rm = TRUE), 2),
    SD = round(sd(Score, na.rm = TRUE), 2),
    Min = round(min(Score, na.rm = TRUE),2),
    Max = round(max(Score, na.rm = TRUE),2),
    Range = Max - Min,
    IQR = round(IQR(Score, na.rm = TRUE),2)
  ) %>%
  ungroup()

# Change the variable names 
program_type_summary <- summary_stats %>%
  rename(
    "Program Type" = program.type,
    "Survey Type" = Survey_Type
  )
```

I will use pivot_wider() to make a more accessible table.

```{r}
#| label: pivot survey scores by program type
#| echo: TRUE
#| results: hide

summary_stats_wide <- summary_stats %>%
  pivot_wider(names_from = program.type, values_from = c(Mean, Median, SD, Min, Max, Range, IQR))%>%
  rename_with(~ gsub("_", " ", .x))
```

```{r}
#| label: tbl-wider-survey-scores-by-program-type
#| tbl-cap: "Survey Scores by Program Type Wide Table"

kable_apa(summary_stats_wide)%>%
  add_header_above(c(" " = 1, "Descriptive Statistics" = 12)) %>%
  column_spec(1, bold = TRUE)

```

From @tbl-wider-survey-scores-by-program-type we can see that for undergraduates, the mean value for Perfectionism Self-Presentation is 5.28, whereas for graduates it's 4.77.

I want to test whether this difference is significant using one-tailed t-test.

My hypotheses are:

Null Hypothesis ($H_0$): Undergraduate's Perfectionism Self-Presentation score does not significantly differ from graduates' score.

Alternative Hypothesis ($H_A$): Undergraduate's Perfectionism Self-Presentation score is significantly higher than graduates' score.

```{r}
#| label: one-tailed t-test
#| echo: true

# Extract scores for each group
undergrad_scores <- math.anxiety %>%
  filter(program.type == "Undergraduate") %>%
  pull(perf.sp)

grad_scores <- math.anxiety %>%
  filter(program.type == "Graduate") %>%
  pull(perf.sp)

# Perform One-Tailed Independent t-Test
t_test_result <- t.test(undergrad_scores, grad_scores, 
                        alternative = "greater", 
                        var.equal = TRUE)  # Assume equal variance

```

```{r}
#| label: tbl-t-test
#| tbl-cap: "One-Tailed t-Test Results for Perfectionism Self-Presentation"
# Extract key statistics from t-test result
t_test_summary <- data.frame(
  Statistic = c("t-value", "df", "p-value", "Mean (Undergraduates)", "Mean (Graduates)", "95% CI Lower", "95% CI Upper"),
  Value = c(
    round(t_test_result$statistic, 3),
    round(t_test_result$parameter, 0),
    format.pval(t_test_result$p.value, digits = 3, eps = .001),
    round(t_test_result$estimate[1], 3),
    round(t_test_result$estimate[2], 3),
    round(t_test_result$conf.int[1], 3),
    round(t_test_result$conf.int[2], 3)
  )
)

kable_apa(t_test_summary) %>%
  column_spec(1, bold = TRUE)


```

As shown in @tbl-t-test, the p-value is `r format.pval(t_test_result$p.value, digits = 3, eps = .001)`, which is significant. Therefore we can reject the null hypothesis and say that undergraduates' Perfectionism Self-Presentation score is significantly higher than graduates' score.

Let's calculate Cohen's d to see the effect size.

```{r}
#| label: cohen's d
#| echo: true

# Define a function to compute Cohen’s d
calculate_cohens_d <- function(x, y) {
  mean_diff <- mean(x, na.rm = TRUE) - mean(y, na.rm = TRUE)
  pooled_sd <- sqrt((sd(x, na.rm = TRUE)^2 + sd(y, na.rm = TRUE)^2) / 2)
  cohens_d <- mean_diff / pooled_sd
  return(cohens_d)
}

cohens_d_value <- calculate_cohens_d(undergrad_scores, grad_scores)

print(paste("Cohen's d:", cohens_d_value))

```

`r paste("Cohen's d =", format(round(cohens_d_value, 3), nsmall = 3))` suggests that the difference in means between undergraduates and graduates is small to moderate in standardized terms.

Now I want to further test whether students' Perfectionism Self-Presentation scores differ by ethnicity.

My hypothesis is that there will be no differences: Null Hypothesis ($H_0$): Students' Perfectionism Self-Presentation scores do not significantly differ by ethnicity.

First, I will set white people as the reference group. Then, I will keep the top three frequent ethinicity and combine the rest as "Other".

```{r}
#| label: anova-set-up
#| echo: TRUE
#| results: hide

# First, I will set white people as the reference group
math.anxiety.ethnicity <- math.anxiety %>%
  mutate(ethnicity.r = fct_relevel(ethnicity.r, "White"))

# Keep the top three frequent ethinicity and combine the rest as "Other"

math.anxiety.ethnicity <- math.anxiety.ethnicity %>%
  mutate(ethnicity.r = fct_lump_n(ethnicity.r, n = 3, other_level = "Other"))


```

Next, I will change the numeric scores of Perfectionism Self-Presentation to three levels: low (score \< 3), medium（3 ≤ score \< 5) and high (5 ≤ score \< 7).

```{r}
#| label: anova-levels
#| echo: TRUE
#| results: hide

math.anxiety.ethnicity <- math.anxiety.ethnicity %>%
  mutate(
    Anxiety_Level = case_when(
      perf.sp < 3 ~ "Low",
      perf.sp >= 3 & perf.sp < 5 ~ "Medium",
      perf.sp >= 5 ~ "High"
    ),
    High_Anxiety = if_else(perf.sp >= 4, "Yes", "No") 
  )

```

Note that I have also created two levels, Perfectionism Self-Presentation ≥4 to be high anxiety and \<4 to be low anxiety. This is independent from the three "low" "medium" "high" anxiety levels. I have to admit that for this mini-project this is purely for the sake of meeting the object of using both "case_when" and "if_else", but in reality one can also perform Chi-square Test using these two levels.

Now I will perform ANOVA and produce APA table.

```{r}
#| label: tbl-anova
#| echo: TRUE


anova_ethnicity <- aov(perf.sp ~ ethnicity.r, data = math.anxiety.ethnicity)

anova_table <- broom::tidy(anova_ethnicity) %>%
  rename(
    `Source` = term,
    `Df` = df,
    `Sum Sq` = sumsq,
    `Mean Sq` = meansq,
    `F value` = statistic,
    `p-value` = p.value
  ) %>%
  mutate(
    `p-value` = ifelse(`p-value` < .001, "< .001", round(`p-value`, 3)) 
  )

  kable_apa(anova_table) %>%
  column_spec(1, bold = TRUE)


```

As shown in Table 4, there was no significant effect of ethnicity, F(`r formatC(summary(anova_ethnicity)[[1]][["F value"]][1], digits = 2, format = "f")`, `r summary(anova_ethnicity)[[1]][["Df"]][1]`, `r summary(anova_ethnicity)[[1]][["Df"]][2]`) = `r formatC(summary(anova_ethnicity)[[1]][["F value"]][1], digits = 2, format = "f")`, p = `r ifelse(summary(anova_ethnicity)[[1]][["Pr(>F)"]][1] < .001, "< .001", formatC(summary(anova_ethnicity)[[1]][["Pr(>F)"]][1], digits = 3, format = "f"))`.

This means the null hypothesis was not rejected and Perfectionism Self-Presentation does not vary across different ethnicity.







# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Appendix

# Title for Appendix
